---
title: "Property Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-deps, message=FALSE, warning=FALSE, echo=F}
#load dependencies
require(plyr)
require(dplyr)
require(osmdata)
require(Hmisc)
require(usedist)
require(leaflet)
require(sf)
require(geosphere)
require(doParallel)
require(foreach)
require(xtable)
require(DT)
```

## Features
`ID`: A unique ID string

`suburb`: The suburb the sale took place in

`add`: Address

`nrooms`: The number of rooms. this may be innacurate, see the report for a discussion

`type`: The type of the property, one of 'House', 'Townhouse' or 'unit'.

`method`: The method of sale: 
S -> sold, SP-> sold prior, PI -> passed in, PN -> sold prior not disclosed, SN -> sold not disclosed, NB -> no bid, VB -> vendor bid, W -> withdrawn prior to auction, SA -> sold after auction, SS -> sold after auction, price not disclosed

`agent`: The Real estate agent

`date`: Date of sale, this appears to be rounded to a period

`precomputeddist`: distance from the CBD in Km

`pc`: postcode

`nbedroom`: number of bedrooms

`nbathroom`: number of bathrooms

`ncar`: number of car parks

`land_area`: Land area in square metres

`building_area`: building area in square metres

`year_built`: year built

`council_area`: The name of the area, named after the governing council

`lat`: latitude

`lng`: longitude

`region`: Greater region name

`propcount`: NUmber of properties in the suburb

`building_area_log`: log of `building_area`

`land_area_log` log(`land_area` + 1)

`price`: sale price

## DataFrame
```{r , results='asis' , echo=F}
load('property_data.Robject')
property_data <- cbind(property_data$ID , select(property_data,-c(ID,price)) , property_data$price)
head(property_data,100) %>% datatable
```

[back to analysis](report.html#data)
